---
title: "p8105_hw4_jh3909"
author: "Jingxuan He"
date: "October 23, 2017"
output: html_document
---

### Package Prepare
```{r, warning = FALSE}
library(tidyverse)
library(haven)
library(readxl)
library(janitor)
library(ggridges)
library(hexbin)
library(knitr)
library(grid)
```

# Read and Clean the Dataï¼Jingxuan
```{r}
# clean the pols data, round the date to fisrt day of the month and rename the mon variable
pols_month = read_csv("./pols-month.csv") %>% 
  clean_names() %>%
  mutate(mon = lubridate::floor_date(mon,"month"))%>%
  rename(date = mon)

# Create a president variable
pols_month = mutate(pols_month, president = prez_dem)
pols_month[pols_month$president == 0, "president"] = "gop"
pols_month[pols_month$president == 1, "president"] = "dem"
pols_month = select(pols_month, -prez_dem, -prez_gop)

#clean the gdp data, round the date to fisrt day of the month
gdp = read_csv("./GDP.csv") %>%
  clean_names() %>%
  mutate(date = lubridate::floor_date(date, "month"), gdp_value = value) %>% # replaced by a more informatic variable name
  filter(!(value == ".")) # remove values with "."

#combine the pols and gdp data
combined_data = left_join(pols_month, gdp, by = "date")
combined_data
```

# Create a plot showing the GDP value over time, including the party of the president in an informative aesthetic- Yue

```{r}
# Make a spaghetti plot

combined_data %>% 
  filter(!is.na(gdp_value)) %>% 
  mutate(gdp_value = as.numeric(gdp_value)) %>% 
  ggplot(., aes(x = date, y = gdp_value,colour=factor(president))) + 
  labs(list(title = "GDP value over time", y = paste("gdp_value"))) + 
  geom_line(aes(group=1)) +
  theme_bw() +
  scale_x_date(date_labels="%b %y",date_breaks  ="36 month")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(title = element_text(size = 20) ,text = element_text(size=15),  
        axis.text.x = element_text(size = 15), 
        legend.text = element_text(size = 15), legend.title = element_text(size = 15)) + 
  scale_colour_discrete(name = "President")
```
```

**Results and explanation**: We made a spaghetti plot to see how the GDP value change over time and grouped the GDP value by the factor that whether the president is republican or democratic. From the graph, we can see that between August 1947 and August 1950, there is a big fluctuation of GDP value when the president was democratic. Additionally, there is peak at around 19 GDP value in 1951. Overall trend of the GDP value remained steady while in 1977, there is a small peak.
